---
title: "Sharpie vs Hurricane Dorian on Twitter"
author: "Joseph Holler"
date: "9/1/2021"
output: html_document
---

# Sharpie vs Hurricane Dorian on Twitter
Author: Joseph Holler
Created: Fall 2019
Updated: Fall 2021

This analysis was developed with assistance from:
- Casey Lilley's GEOG 323 final project available at: https://caseylilley.github.io/finalproj.html  
- Leah Wasser and Carson Farmer's *Twitter Data in R Using RTweet* 
tutorial on EarthLab at: 
https://www.earthdatascience.org/courses/earth-analytics/get-data-using-apis/use-twitter-api-r/
- Michael Minn's *Basic Spatial Point Pattern Analysis in R* tutorial at: http://michaelminn.net/tutorials/r-point-analysis/

## Set up environment

Load the R project saved in the root directory of this repository, so that the
working directory is the root directory of the repository.

```{r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)

# list of required packages
packages <- c(
  "here", "svDialogs", "tidyverse",
  "rtweet", "rehydratoR",
  "tidytext", "tm", "igraph", "ggraph",
  "tidycensus", "sf", "spdep"
)

# additional packages: test which are still needed:
# "maps", "tm", "RColorBrewer", "rccmisc",

# load and install required packages
package_check <- lapply(
  packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE, quietly = TRUE)
      library(x, character.only = TRUE)
    }
  }
)
```

```{r save R processing environment}
# save the R processing environment, including date in file name
writeLines(
  capture.output(sessionInfo()),
  here("procedure", "environment", paste0("r-environment-", Sys.Date(), ".txt"))
  )
```


## Set up Twitter Application

You will need Twitter Developer API access to run this analysis:
See https://cran.r-project.org/web/packages/rtweet/vignettes/auth.html for
instructions.

reference for search_tweets function: 
https://rtweet.info/reference/search_tweets.html 
- don't add any spaces in between variable name and value for your search  
e.g. n=1000 is better than n = 1000
- the first parameter in quotes is the search string
- `n=10000` asks for 10,000 tweets
- if you want more than 18,000 tweets, change `retryonratelimit` to `TRUE` and 
wait 15 minutes for every batch of 18,000
- `include_rts=FALSE` excludes retweets.
- `token` refers to the twitter token you defined above for access to your twitter
developer account
- `geocode` is equal to a string with three parts: longitude, latitude, and 
distance with the units mi for miles or km for kilometers

This code block will ask you for your twitter application name, key, and secret.
Then it will launch a web browser and prompt you to log in to Twitter to 
authenticate the application.

Never save your API keys in code where it can be committed and synced to GitHub!
The code below is configured to save your keys in the environment, and this Git
repository is set up to ignore the R environment data file.

```{r twitter-setup, eval = FALSE}
# Twitter application values
twitter_vars <- list(
  app = "enter Twitter application name",
  key = "enter Twitter API key",
  secret = "enter Twitter API secret key"
)

# if Twitter token has already been created, auto-fill dialogue with its values
if (exists("twitter_token")) {
  twitter_vars$app <- twitter_token$app$appname
  twitter_vars$key <- twitter_token$app$key
  twitter_vars$secret <- twitter_token$app$secret
}

twitter_token <- create_token(
  app = dlgInput("Twitter App Name:", twitter_vars$app)$res,
  consumer_key = dlgInput("Consumer Key:", twitter_vars$key)$res,
  consumer_secret = dlgInput("Consumer Secret:", twitter_vars$secret)$res,
  access_token = NULL,
  access_secret = NULL
)
```

# Pre-processing

- Acquire Twitter data for analysis
- Filter Twitter data for good geographic information and convert to Lat/Long
coordinates

## Search for Hurricane Dorian tweets

get tweets for hurricane Dorian, searched on September 11, 2019
**Warning**: this code will no longer result in the same data! 
It is here for reference or replication work only.

```{r search-dorian, eval = FALSE}
dorian_raw <- search_tweets("dorian OR hurricane OR sharpiegate",
  n = 200000, include_rts = FALSE,
  token = twitter_token,
  geocode = "32,-78,1000mi",
  retryonratelimit = TRUE
)

# write status id's for results of the original twitter search
write.table(dorian_raw$status_id,
  here("data", "raw", "public", "dorianids.txt"),
  append = FALSE, quote = FALSE, row.names = FALSE, col.names = FALSE
)
```

## Search for generic tweets after hurricane season

get tweets without any text filter for the same geographic region in November, 
searched on November 19, 2019
the query searches for all verified or unverified tweets, i.e. everything

**Warning**: this code will no longer result in the same data! 
It is here for reference or replication work only.

```{r search-november, eval = FALSE}
november_raw <- search_tweets("-filter:verified OR filter:verified",
  n = 200000, include_rts = FALSE,
  token = twitter_token,
  geocode = "32,-78,1000mi",
  retryonratelimit = TRUE
)

# write status id's for results of the original twitter search
write.table(november_raw$status_id,
  here("data", "raw", "public", "novemberids.txt"),
  append = FALSE, quote = FALSE, row.names = FALSE, col.names = FALSE
)
```

## Rehydrate the twitter data

Twitter does not permit redistribution of Twitter data, with the exception of 
tweet status ids. 
For the purposes of transparency and reproducibility, 
researchers may include a list of status id's with their publication. 
The process of using those status ids to query Twitter for the full tweet data 
is called **rehydrating**&mdash;like going back to the analogous fire hose of 
big data.
**Warning**: Twitter users and individual tweets can be deleted over time, 
therefore the results of rehydration will be similar, but not identical to, the
original Twitter data used for this research.  
**Warning**: It may take more than an hour to rehydrate the raw tweets with 
Twitter queries, therefore you may select to load only the derived status ids,
which have filtered only the tweets with valid geographic data (approximately 
one tenth of the raw tweets)

### Load Twitter status ids

You have a choice of loading all of the status IDs from the original search, or a list of status IDs filtered to contain only those with precise enough geographic data.
Loading the original search will be a more exact replication while loading filtered data will speed up processing without changing the final result.
Neither process will result in identical findings compared to the original study because Tweets can be deleted after posting.

```{r load-original-statusids, eval = FALSE}
# load tweet status id's for Hurricane Dorian search results

filtered <- dlgList(
  choices = c("raw", "derived"), 
  title = "Which Dorian ids?")$res

dorianids <-
  data.frame(read.table(here("data", filtered, "public", "dorianids.txt"),
    numerals = "no.loss"
  ))

# load cleaned status id's for November general twitter search
filtered <- dlgList(
  choices = c("raw", "derived"), 
  title = "Which November ids?")$res


novemberids <-
  data.frame(read.table(here("data", "raw", "public", "novemberids.txt"),
    numerals = "no.loss"
  ))
```

### Rehydrate Twitter status ids
This operation may take over an hour to run on all of the raw tweets

```{r rehydrate, eval = FALSE}
# rehydrate dorian tweets
dorian_raw <- rehydratoR(twitter_token$app$key, twitter_token$app$secret,
  twitter_token$credentials$oauth_token,
  twitter_token$credentials$oauth_secret, dorianids,
  base_path = NULL, group_start = 1
)

# rehydrate november tweets
november_raw <- rehydratoR(twitter_token$app$key, twitter_token$app$secret,
  twitter_token$credentials$oauth_token,
  twitter_token$credentials$oauth_secret, novemberids,
  base_path = NULL, group_start = 1
)
``` 

## Load the original search results

Students in the GEOG 323 Open Source GIScience course may download the original
search results from the private course data repository: 
https://github.com/GIS4DEV/geog323data/raw/main/dorian/dorian_raw.RDS
https://github.com/GIS4DEV/geog323data/raw/main/dorian/november.RDS

Save the two `.RDS` files to the `data/raw/private` folder and then load the
data with the code block below.

```{r load-original}
dorian_raw <- readRDS(here("data", "raw", "private", "dorian_raw.RDS"))
november_raw <- readRDS(here("data", "raw", "private", "november_raw.RDS"))
```

## Process geographic data in tweets 

reference for lat_lng function: https://rtweet.info/reference/lat_lng.html
adds a lat and long field to the data frame, picked out of the fields
that you indicate in the c() list
sample function: lat_lng(x, coords = c("coords_coords", "bbox_coords"))

list and count unique place types
NA results included based on profile locations, not geotagging / geocoding.
If you have these, it indicates that you exhausted the more precise tweets 
in your search parameters and are including locations based on user profiles

```{r count-place-types}
count(dorian_raw, place_type)
```

### Convert geographic information into lat/long coordinates

If you have loaded filtered status ids, or you have already run this code, you
will not notice a difference in `place_type` or `n` because the data has already
been processed.

```{r lat-long}
# convert GPS coordinates into lat and lng columns
# do not use geo_coords! Lat/Lng will be inverted
dorian <- lat_lng(dorian_raw, coords = c("coords_coords"))
november <- lat_lng(november_raw, coords = c("coords_coords"))

# select any tweets with lat and lng columns (from GPS) or
# designated place types of your choosing
dorian <- subset(
  dorian,
  place_type == "city" | place_type == "neighborhood" |
    place_type == "poi" | !is.na(lat)
)

november <- subset(
  november,
  place_type == "city" | place_type == "neighborhood" |
    place_type == "poi" | !is.na(lat)
)

# convert bounding boxes into centroids for lat and lng columns
dorian <- lat_lng(dorian, coords = c("bbox_coords"))
november <- lat_lng(november, coords = c("bbox_coords"))

# re-check counts of place types
count(dorian, place_type)
```

### Save processed tweets

Optionally,
Save the tweet id's to the `data\derived\public` folder as plain text.  
Save the full tweet data to `data\derived\private` folder as RDS files.  
Full Tweet data cannot be shared with the public, therefore it is stored in
a folder ignored by Git.

```{r save-processed-tweets}
write.table(november$status_id,
  here("data", "derived", "public", "novemberids.txt"),
  append = FALSE, quote = FALSE, row.names = FALSE, col.names = FALSE
)

write.table(dorian$status_id,
  here("data", "derived", "public", "dorianids.txt"),
  append = FALSE, quote = FALSE, row.names = FALSE, col.names = FALSE
)

saveRDS(dorian, here("data", "derived", "private", "dorian.RDS"))
saveRDS(november, here("data", "derived", "private", "november.RDS"))
```

### Load processed tweets

Optionally,
Load processed twitter data

```{r load-processed-tweets, eval = FALSE }
dorian <- readRDS(here("data", "derived", "private", "dorian.RDS"))
november <- readRDS(here("data", "derived", "private", "november.RDS"))
```

# Temporal Analysis

Create a temporal data frame and graph it

```{r temporal}
dorian_tweets_by_hour <- ts_data(dorian, by = "hours")
ts_plot(dorian, by = "hours")
```

# Network Analysis

Create a network data frame. 
Other options for 'edges' in the network include mention, retweet, and reply
The graph does not look good (understatement), because of poor data regarding quotes.
It would be more interesting to visualize retweets, but we did not search for
them, as we were focusing on original content.

```{r network}
dorian_network <- network_graph(dorian, c("quote"))
plot.igraph(dorian_network)
```

# Text Analysis

## Clean the text data

Parse the tweet data for plain language, and parse tweet text into words.
Remove stop words and our search terms.

```{r text-processing}
# remove urls, fancy formatting, etc. in other words, clean the text content
dorian_text <- dorian %>%
  select(text) %>%
  plain_tweets()

# parse out words from tweet text
dorian_words <- dorian_text %>% unnest_tokens(word, text)

# how many words do you have including the stop words?
word_count <- list(before = count(dorian_words)$n)

# create list of stop words (useless words not worth analyzing)
data("stop_words")

# add "t.co" twitter links to the list of stop words
# also add the twitter search terms to the list
stop_words <- stop_words %>%
  add_row(word = "t.co", lexicon = "SMART") %>%
  add_row(word = "hurricane", lexicon = "Search") %>%
  add_row(word = "dorian", lexicon = "Search") %>%
  add_row(word = "sharpiegate", lexicon = "Search")

# delete stop words from dorianWords with an anti_join
dorian_words <- anti_join(dorian_words, stop_words, by="word")

# how many words after removing the stop words?
word_count <- append(
  word_count,
  list(after = count(dorian_words)$n)
  )
print(word_count)
```

## Graph frequencies of words

```{r word-frequency}
dorian_words %>%
  count(word, sort = TRUE) %>%
  top_n(15) %>%  #deprecated; try replacing with slice_min() or slice_max()
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(x = word, y = n)) +
  geom_col() +
  xlab(NULL) +
  coord_flip() +
  labs(
    x = "Count",
    y = "Unique words",
    title = "Count of unique words found in tweets"
  )
```

## Analyze and graph word assoctiation

```{r word-association}
# separate words and count frequency of word pair occurrence in tweets
dorian_word_pairs <- dorian_text %>%
  mutate(text = removeWords(tolower(text), stop_words$word)) %>%
  unnest_tokens(paired_words, text, token = "ngrams", n = 2) %>%
  separate(paired_words, c("word1", "word2"), sep = " ") %>%
  count(word1, word2, sort = TRUE)

# graph a word cloud with space indicating association.
# you may change the filter to filter more or less than pairs with 25 instances
dorian_word_pairs %>%
  filter(n >= 25 & !is.na(word1) & !is.na(word2)) %>%
  graph_from_data_frame() %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(edge_alpha = n, edge_width = n)) +
  geom_node_point(color = "darkslategray4", size = 3) +
  geom_node_text(aes(label = name), vjust = 1.8, size = 3) +
  labs(
    title = "Word Network of Tweets during Hurricane Dorian",
    x = "", y = ""
  ) +
  theme(
    plot.background = element_rect(
    fill = "grey95",
    colour = "black",
    size = 1
    ),
    legend.background = element_rect(fill = "grey95")
  )
```

## Spatial Analysis

First, you will need a Census API. You can sign up for one here: https://api.census.gov/data/key_signup.html

```{r census-api}
census_api_key(dlgInput(
  "Enter a Census API Key",
  Sys.getenv("CENSUS_API_KEY")
)$res,
overwrite = TRUE
)

counties <- get_estimates(
  "county",
  product = "population",
  output = "wide",
  geometry = TRUE,
  keep_geo_vars = TRUE
)
```

## Select and save counties of interest

select only the states you want, with FIPS state codes
look up fips codes here: 
https://en.wikipedia.org/wiki/Federal_Information_Processing_Standard_state_code 

```{r save-counties}
counties <- filter(
  counties,
  STATEFP %in% c(
    "54", "51", "50", "47", "45", "44", "42", "39", "37", "36", "05", "01",
    "34", "33", "29", "28", "25", "24", "23", "22", "21", "18", "17", "13",
    "12", "11", "10", "09"
  )
)
# alternatively, one could select by location using original search criteria...

saveRDS(counties, here("data", "derived", "public", "counties.RDS"))
```

### Load counties

Optionally, load counties from the `counties.RDS` file saved with the repository

```{r load-counties}
counties <- readRDS(here("data", "derived", "public", "counties.RDS"))
```

## Map Population Density and Tweet Points

map results with GGPlot
note: cut_interval is an equal interval classification function, while 
cut_number is a quantile / equal count function
you can change the colors, titles, and transparency of points

```{r map-tweet-points}
ggplot() +
  geom_sf(data = counties, 
    aes(fill = cut_number(DENSITY, 5)), color = "grey") +
  scale_fill_brewer(palette = "GnBu") +
  guides(fill = guide_legend(title = "Population Density")) +
  geom_point(
    data = dorian, aes(x = lng, y = lat),
    colour = "purple", alpha = 0.1, size = 1
  ) +
  labs(title = "Tweet Locations During Hurricane Dorian") +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  ) +
  xlim(-96, -67) +
  ylim(24, 47)
```

## Join Tweets to Counties

Spatially join Dorian and November tweets to counties and save counties
as `counties_tweet_counts.RDS`

```{r join-tweets-to-counties}
dorian_sf <- dorian %>%
  st_as_sf(coords = c("lng", "lat"), crs = 4326) %>% # make point geometries
  st_transform(4269) %>% # transform to NAD 1983
  st_join(select(counties, GEOID)) # spatially join counties to each tweet

dorian_by_county <- dorian_sf %>%
  st_drop_geometry() %>% # drop geometry / make simple table
  group_by(GEOID) %>% # group by county using GEOID
  summarise(dorian = n()) # count # of tweets

counties <- counties %>%
  left_join(dorian_by_county, by = "GEOID") %>% # join count of tweets to counties
  mutate(dorian = replace_na(dorian, 0)) # replace nulls with 0's

rm(dorian_by_county)

# Repeat the workflow above for tweets in November

nov_by_county <- november %>%
  st_as_sf(coords = c("lng", "lat"), crs = 4326) %>%
  st_transform(4269) %>%
  st_join(select(counties, GEOID)) %>%
  st_drop_geometry() %>%
  group_by(GEOID) %>%
  summarise(nov = n())

counties <- counties %>%
  left_join(nov_by_county, by = "GEOID") %>%
  mutate(nov = replace_na(nov, 0))

counties <- counties %>%
  mutate(dorrate = dorian / POP * 10000) %>% # dorrate is tweets per 10,000
  mutate(ndti = (dorian - nov) / (dorian + nov)) %>% # normalized diff tweet index
  mutate(ndti = replace_na(ndti, 0)) # replace NULLs with 0's

rm(nov_by_county)

# save counties geographic data with derived tweet rates
saveRDS(counties, here("data", "derived", "public", "counties_tweet_counts.RDS"))
```

Optionally, begin here by loading counties with Twitter data

```{r load-counties-with-twitter}
counties <- readRDS(here("data", "derived", "public", "counties_tweet_counts.RDS"))
```

```{r map ndti}
ggplot() +
  geom_sf(data = counties, 
    aes(fill = cut_interval(dorrate, 5)), color = "grey", lwd=0.01) +
  scale_fill_brewer(palette = "GnBu") +
  guides(fill = guide_legend(title = "Normalized Difference Tweet Index")) +
  labs(title = "Hurricane Dorian Twitter Activity") +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  ) +
  xlim(-96, -67) +
  ylim(24, 47)
```

## Spatial Cluster Analysis

### Create Spatial Weight Matrix

Use 110km Euclidean distance and include self in the weight matrix

```{r spatial-weight-matrix, warning = FALSE}
county_coords <- counties %>%
  st_centroid() %>% # convert polygons to centroid points
  st_coordinates() # convert to simple x,y coordinates to play with stdep

thresdist <- county_coords %>% 
  dnearneigh(0, 110, longlat = TRUE) %>% # use geodesic distance of 110km
  # distance should be long enough for every feature to have >= one neighbor
  include.self() # include a county in its own neighborhood (for G*)

thresdist # view statistical summary of the nearest neighbors
```

Optionally, plot the spatial weight matrix results
This should result in a very dense graph, because each county is connected
to all other counties within 110 km.

```{r plot-spatial-weight-matrix}
plot(thresdist, county_coords, lwd=0.01) # plot nearest neighbor ties
```

## Calculate Getis-Ord G* Statistic

```{r getis-ord}

# Create weight matrix from the neighbor objects
dwm <- nb2listw(thresdist, zero.policy = T)

######## Local G* Hotspot Analysis ########
# Get Ord G* statistic for hot and cold spots
counties$locG <- as.vector(localG(counties$dorrate,
  listw = dwm,
  zero.policy = TRUE
))

# may be same as:
# counties$dorrate %>% localG(listw = dwm, zero.policy = TRUE) %>% as.vector()

# check summary statistics of the local G score
summary(counties$locG)
```

## Map Hotpots

classify G scores by significance values typical of Z-scores
where 1.15 is at the 0.125 confidence level,
and 1.95 is at the 0.05 confidence level for two tailed z-scores
based on Getis and Ord (1995) Doi: 10.1111/j.1538-4632.1992.tb00261.x
to find other critical values, use the qnorm() function as shown here:
https://methodenlehre.github.io/SGSCLM-R-course/statistical-distributions.html
Getis Ord also suggest applying a Bonferroni correction 

breaks and colors from http://michaelminn.net/tutorials/r-point-analysis/
based on 1.96 as the 95% confidence interval for z-scores
if your results don't have values in each of the 5 categories, you may need
to change the values & labels accordingly.

Can we figure out a way to include the neighborhood surrounding each significant
cluster?

```{r map-hotspots}
# classify by significance levels
siglevel <- c(1.15, 1.95)
counties <- counties %>%
  mutate(sig = cut(locG, c(
    min(counties$locG),
    siglevel[2] * -1,
    siglevel[1] * -1,
    siglevel[1],
    siglevel[2],
    max(counties$locG)
  )))
rm(siglevel)

# map results!
ggplot() +
  geom_sf(data = counties, aes(fill = sig), color = "white", lwd = 0.1) +
  scale_fill_manual(
    values = c("#0000FF80", "#8080FF80", "#FFFFFF80", "#FF808080", "#FF000080"),
    labels = c("low", "", "insignificant", "", "high"),
    aesthetics = "fill"
  ) +
  guides(fill = guide_legend(title = "Hot Spots")) +
  labs(title = "Clusters of Hurricane Dorian Twitter Activity") +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  )
```
